---
title: "Contrast Algorithms for Linear Models"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: default
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Using Common Contrast Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
library(CALM)
```

## Using Common Contrast Codes

### The Data

For the sake of example, this site uses the following data. For reference, the group means are displayed.

```{r}
mtcars
mtcars$cyl <- factor(mtcars$cyl) 
tapply(mtcars$mpg,mtcars$cyl,mean)
```
### Treatment (First) Contrasts

```{r}
contrasts(mtcars$cyl) <- treatment.first
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the first group. Parameters represent the difference between each group and the first group.

### Treatment (Last) Contrasts

```{r}
contrasts(mtcars$cyl) <- treatment.last
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the last group. Parameters represent the difference between each group and the last group.

### Control (First) Contrasts

```{r}
contrasts(mtcars$cyl) <- control.first
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the first group.

### Control (Last) Contrasts

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the last group.

```{r}
contrasts(mtcars$cyl) <- control.last
summary(lm(mpg~cyl, mtcars))$coefficients
```

### Deviation (First) Contrasts

```{r}
contrasts(mtcars$cyl) <- deviation.first
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the grand mean. The first group is not compared. 

### Deviation (Last) Contrasts

```{r}
contrasts(mtcars$cyl) <- deviation.last
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the grand mean. The last group is not compared.

### Helmert (Forward) Contrasts

```{r}
contrasts(mtcars$cyl) <- helmert.forward
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the mean of the subsequent groups. 

### Helmert (Reverse) Contrasts

```{r}
contrasts(mtcars$cyl) <- helmert.reverse
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the mean of the previous groups.

### Repeated (Forward) Contrasts

```{r}
contrasts(mtcars$cyl) <- repeated.forward
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the first group. Parameters represent the difference between each group and the subsequent group.

### Repeated (Reverse) Contrasts

```{r}
contrasts(mtcars$cyl) <- repeated.reverse
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the first group. Parameters represent the difference between each group and the previous group.

### Difference (Forward) Contrasts

```{r}
contrasts(mtcars$cyl) <- difference.forward
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the subsequent group. 

### Difference (Reverse) Contrasts

```{r}
contrasts(mtcars$cyl) <- difference.reverse
summary(lm(mpg~cyl, mtcars))$coefficients
```

The intercept is equivalent to the mean of the group means. Parameters represent the difference between each group and the previous group. 
