---
title: "Contrast Algorithms for Linear Models"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: default
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Extracting Hypotheses from Regression Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
library(CALM)
```

## Extracting from Regression Codes

**CALM** can take existing coding patterns and decode them to determine the hypotheses being tested. The examples below use traditional and commonly-taught regression codes (dummy codes, effect codes, etc.) and extract the hypotheses implied by each code vector.

The package takes matrix input of the coding pattern (columns represent coded vectors, rows represent the groups being represented) and outputs the linear contrasts hypotheses being represented (the rows represent the weights for the contrast, the columns represent the groups being compared).

### Dummy Coding

```{r}
ex_dummy <- matrix(c(
    1,1,0,
    1,0,1,
    1,0,0),
    ncol=3,
    byrow=TRUE)
calm.decode(ex_dummy)
```

The intercept is equivalent to the mean of the last group. Parameters represent the difference between each group and the last group. 

### Effect Coding

```{r}
ex_effect <- matrix(c(
    1,1,0,
    1,0,1,
    1,-1,-1),
    ncol=3,
    byrow=TRUE)
calm.decode(ex_effect)
```

The intercept is equivalent to the mean of the group means. Parameter represent the difference between each group and the grand mean. The last group is not compared. 

### Repeated Coding

```{r}
ex_repeated <- matrix(c(
    1,0,0,
    1,1,0,
    1,1,1),
    ncol=3,
    byrow=TRUE)
calm.decode(ex_repeated)
```

The intercept is equivalent to the mean of the first group. Parameter represent the difference between each group and the previous group. 
