---
title: "Contrast Algorithms for Linear Models"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: default
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Using Common Contrast Matrices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
library(CALM)
```

## Using Common Contrast Matrices

### Overview



### Example: Treatment (First) Contrasts

```{r}
treatment.first(3)
treatment.first(3,intercept=TRUE)
calm.decode(treatment.first(3))
calm.decode(treatment.first(3,intercept=TRUE))
```

### Example: Treatment (Last) Contrasts

```{r}
treatment.last(3)
treatment.last(3,intercept=TRUE)
calm.decode(treatment.last(3))
calm.decode(treatment.last(3,intercept=TRUE))
```

### Example: Deviation (First) Contrasts

```{r}
deviation.first(3)
deviation.first(3,intercept=TRUE)
calm.decode(deviation.first(3))
calm.decode(deviation.first(3,intercept=TRUE))
```

### Example: Deviation (Last) Contrasts

```{r}
deviation.last(3)
deviation.last(3,intercept=TRUE)
calm.decode(deviation.last(3))
calm.decode(deviation.last(3,intercept=TRUE))
```

### Example: Helmert (Forward) Contrasts

```{r}
helmert.forward(3)
helmert.forward(3,intercept=TRUE)
calm.decode(helmert.forward(3))
calm.decode(helmert.forward(3,intercept=TRUE))
```

### Example: Helmert (Reverse) Contrasts

```{r}
helmert.reverse(3)
helmert.reverse(3,intercept=TRUE)
calm.decode(helmert.reverse(3))
calm.decode(helmert.reverse(3,intercept=TRUE))
```

### Example: Repeated (Forward) Contrasts

```{r}
repeated.forward(3)
repeated.forward(3,intercept=TRUE)
calm.decode(repeated.forward(3))
calm.decode(repeated.forward(3,intercept=TRUE))
```

### Example: Repeated (Reverse) Contrasts

```{r}
repeated.reverse(3)
repeated.reverse(3,intercept=TRUE)
calm.decode(repeated.reverse(3))
calm.decode(repeated.reverse(3,intercept=TRUE))
```

### References

Wendorf, C. A. (2004). Primer on multiple regression coding: Common forms and the additional case of repeated contrasts. *Understanding Statistics, 3,* 47-57. 
