---
subtitle: "Extracting from R BAse Contrast Codes"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Extracting from R Base Contrast Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7",filename='revised_toc.R')
```

```{r,include=FALSE}
if(!require(CALM)) {
if(!require(remotes)) install.packages('remotes')
remotes::install_github("cwendorf/CALM")}
library(CALM)
```

## Extracting Hypotheses from Base R Contrast Codes

**CALM** can take existing coding patterns and decode them to determine the hypotheses being tested. The examples below use the built-in codes in R (treatment, SAS, sum, helmert, and poly) and extract the hypotheses implied by each code vector.

The package takes matrix input of the coding pattern (columns represent coded vectors, rows represent the groups being represented) and outputs the linear contrasts hypotheses being represented (the rows represent the weights for the contrast, the columns represent the groups being compared).

```{r toc, echo=FALSE}
thisfile <- knitr::current_input()
revised_toc(thisfile,base_level=3,toc_depth=4)
```

---

### contr.treatment

Specify that we wish to examine the existing `contr.treatment` and then use CALM to decode it into hypotheses.

```{r}
ex_contr.treatment <- cbind(Int=1,contr.treatment(3))
ex_contr.treatment
calm.decode(ex_contr.treatment)
```

This shows that the intercept is equivalent to the mean of the first group. The additional parameters represent the difference between each group and the first group. 

### contr.SAS

Specify that we wish to examine the existing `contr.SAS` and then use CALM to decode it into hypotheses.

```{r}
ex_contr.SAS <- cbind(Int=1,contr.SAS(3))
ex_contr.SAS
calm.decode(ex_contr.SAS)
```

This shows that the intercept is equivalent to the mean of the last group. The additional parameters represent the difference between each group and the last group.

### contr.sum

Specify that we wish to examine the existing `contr.sum` and then use CALM to decode it into hypotheses.

```{r}
ex_contr.sum <- cbind(Int=1,contr.sum(3))
ex_contr.sum
calm.decode(ex_contr.sum)
```

This shows that the intercept is equivalent to the mean of the group means. The additional parameters represent the difference between each group and the grand mean. The last group is not compared.

### contr.helmert

Specify that we wish to examine the existing `contr.herlmert` and then use CALM to decode it into hypotheses.

```{r}
ex_contr.helmert <- cbind(Int=1,contr.helmert(3))
ex_contr.helmert
calm.decode(ex_contr.helmert)
```

This shows that the intercept is equivalent to the mean of the group means. The additional parameters represent the difference between each group and the mean of the subsequent groups. 

### contr.poly

Specify that we wish to examine the existing `contr.poly` and then use CALM to decode it into hypotheses.

```{r}
ex_contr.poly <- cbind(Int=1,contr.poly(3))
ex_contr.poly
calm.decode(ex_contr.poly)
```

THis shows that the intercept is equivalent to the mean of the group means. The additional parameters represent polynomial trends (linear, quadratic, etc.).
