---
title: "Contrast Algorithms for Linear Models"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: default
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Producing Common Contrast Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
if(!require(CALM)) {
if(!require(remotes)) install.packages('remotes')
remotes::install_github("cwendorf/CALM")}
library(CALM)
```

## Producing Common Contrast Codes

**CALM** can take a set of linear contrasts to be tested and encode them into coding schemes. The examples below take some common contrasts (treatment, control, deviation, etc.) and produce the coding scheme associated with it.

The package takes linear contrast weights (rows represented the weights for the groups, columns represent the different groups) and provides the coding pattern (where columns represent vector codes, rows represent the groups).

### Treatment (First) Contrasts

The intercept is equivalent to mean of the first group. Parameters represent the difference between each group and the first group. 

```{r}
M1 <- c(1,0,0)
H1 <- c(-1,1,0)
H2 <- c(-1,0,1)
ex_treatment.first <- rbind(M1,H1,H2)
calm.encode(ex_treatment.first)
```

### Treatment (Last) Contrasts

The intercept is equivalent to mean of the last group. Parameters represent the difference between each group and the last group. 

```{r}
M3 <- c(0,0,1)
H1 <- c(1,0,-1)
H2 <- c(0,1,-1)
ex_treatment.last <- rbind(M3,H1,H2)
calm.encode(ex_treatment.last)
```

### Control (First) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the first group. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(-1,1,0)
H2 <- c(-1,0,1)
ex_control.first <- rbind(GM,H1,H2)
calm.encode(ex_control.first)
```

### Control (Last) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the last group. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(1,0,-1)
H2 <- c(0,1,-1)
ex_control.last <- rbind(GM,H1,H2)
calm.encode(ex_control.last)
```

### Deviation (First) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the grand mean. The first group is not compared. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(-1/3,2/3,-1/3)
H2 <- c(-1/3,-1/3,2/3)
ex_deviation.first <- rbind(GM,H1,H2)
calm.encode(ex_deviation.first)
```

### Deviation (Last) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the grand mean. The last group is not compared. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(2/3,-1/3,-1/3)
H2 <- c(-1/3,2/3,-1/3)
ex_deviation.last <- rbind(GM,H1,H2)
calm.encode(ex_deviation.last)
```

### Helmert (Forward) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the mean of the subsequent groups. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(1,-1/2,-1/2)
H2 <- c(0,1,-1)
ex_helmert.forward <- rbind(GM,H1,H2)
calm.encode(ex_helmert.forward)
```

### Helmert (Reverse) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the mean of the previous groups. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(-1,1,0)
H2 <- c(-1/2,-1/2,1)
ex_helmert.reverse <- rbind(GM,H1,H2)
calm.encode(ex_helmert.reverse)
```

### Repeated (Forward) Contrasts

The intercept is equivalent to mean of the first group. Parameters represent the difference between each group and the subsequent group. 

```{r}
M1 <- c(1,0,0)
H1 <- c(1,-1,0)
H2 <- c(0,1,-1)
ex_repeated.forward <- rbind(M1,H1,H2)
calm.encode(ex_repeated.forward)
```

### Repeated (Reverse) Contrasts

The intercept is equivalent to mean of the first group. Parameters represent the difference between each group and the previous group. 

```{r}
M1 <- c(1,0,0)
H1 <- c(-1,1,0)
H2 <- c(0,-1,1)
ex_repeated.reverse <- rbind(M1,H1,H2)
calm.encode(ex_repeated.reverse)
```

### Difference (Forward) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the subsequent group. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(1,-1,0)
H2 <- c(0,1,-1)
ex_difference.forward <- rbind(GM,H1,H2)
calm.encode(ex_difference.forward)
```

### Difference (Reverse) Contrasts

The intercept is equivalent to mean of the group means. Parameters represent the difference between each group and the previous group. 

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(-1,1,0)
H2 <- c(0,-1,1)
ex_difference.reverse <- rbind(GM,H1,H2)
calm.encode(ex_difference.reverse)
```
