---
subtitle: "Producing Custom Contrast Codes"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Producing Custom Contrast Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7",filename='revised_toc.R')
```

```{r,include=FALSE}
if(!require(CALM)) {
if(!require(remotes)) install.packages('remotes')
remotes::install_github("cwendorf/CALM")}
library(CALM)
```

## Producing Custom Contrast Codes

**CALM** can take a set of linear contrasts to be tested and encode them into coding schemes. The examples below take custom contrasts (both orthogonal and nonorthogonal) and produce the coding scheme associated with it.

The package takes linear contrast weights (rows represented the weights for the groups, columns represent the different groups) and provides the coding pattern (where columns represent vector codes, rows represent the groups).

```{r toc, echo=FALSE}
thisfile <- knitr::current_input()
revised_toc(thisfile,base_level=3,toc_depth=4)
```

---

### Orthogonal Contrasts

In this custom example, the intercept is equivalent to mean of the group means. The first parameter represents the difference between the first group and mean of the subsequent groups. The second parameter represents the difference between the second and third groups.

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(1,-1/2,-1/2)
H2 <- c(0,1,-1)
ex_custom.orthogonal <- rbind(GM,H1,H2)
calm.encode(ex_custom.orthogonal)
```

### Nonorthogonal Contrasts

In this custom example, the intercept is equivalent to mean of the first group. The first parameter represents the difference between the first group and mean of the subsequent groups. The second parameter represents the difference between the first and second groups.

```{r}
GM <- c(1,0,0)
H1 <- c(1,-1/2,-1/2)
H2 <- c(1,-1,0)
ex_custom.nonorthogonal <- rbind(GM,H1,H2)
calm.encode(ex_custom.nonorthogonal)
```
