---
subtitle: "using Common Contrast Codes"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Using Custom Contrast Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
# suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width = 7, fig.height = 5, fig.path = "figures/", echo = TRUE, warning = FALSE, message = FALSE)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7", filename = "revised_toc.R")
```

```{r,include=FALSE}
if (!require(CALM)) {
  if (!require(remotes)) install.packages("remotes")
  remotes::install_github("cwendorf/CALM")
}
library(CALM)
```

## Using Custom Contrast Codes

`CALM` also makes it possible to develop custom codes to test specific hypotheses. The examples below use custom codes in the context of a General Linear Model to test the associated hypotheses.

```{r toc, echo=FALSE}
thisfile <- knitr::current_input()
revised_toc(thisfile, base_level = 3, toc_depth = 4)
```

---

### The Data

For the sake of example, this site uses the following data. For reference, the group means are displayed.

```{r}
mtcars
mtcars$cyl <- factor(mtcars$cyl)
summary(lm(mpg ~ cyl, data = mtcars))
tapply(mtcars$mpg, mtcars$cyl, mean)
```

### Orthogonal Contrasts

Here, the custom hypotheses a specified and then `CALM` is used to create the proper codes.

```{r}
GM <- c(1 / 3, 1 / 3, 1 / 3)
H1 <- c(1, -1 / 2, -1 / 2)
H2 <- c(0, 1, -1)
ex_custom.orthogonal <- rbind(GM, H1, H2)
custom.orthogonal <- calm.encode(ex_custom.orthogonal)
custom.orthogonal
```

```{r}
contrasts(mtcars$cyl) <- custom.orthogonal[, -1]
summary(lm(mpg ~ cyl, data = mtcars))$coefficients
```

In this custom example, the intercept is equivalent to the mean of the group means. The first parameter represents the difference between the first group and mean of the subsequent groups. The second parameter represents the difference between the second and third groups.

### Nonorthogonal Contrasts

Again, the custom hypotheses a specified and then `CALM` is used to create the proper codes.

```{r}
GM <- c(1, 0, 0)
H1 <- c(1, -1 / 2, -1 / 2)
H2 <- c(1, -1, 0)
ex_custom.nonorthogonal <- rbind(GM, H1, H2)
custom.nonorthogonal <- calm.encode(ex_custom.nonorthogonal)
custom.nonorthogonal
```

```{r}
contrasts(mtcars$cyl) <- custom.nonorthogonal[, -1]
summary(lm(mpg ~ cyl, data = mtcars))$coefficients
```

In this custom example, the intercept is equivalent to the mean of the first group. The first parameter represents the difference between the first group and mean of the subsequent groups. The second parameter represents the difference between the first and second groups.
