---
title: "Contrast Algorithms for Linear Models"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: default
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Using Custom Contrast Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
library(CALM)
```

## Using Custom Contrast Codes

### The Data

For the sake of example, this site uses the following data. For reference, the group means are displayed.

```{r}
mtcars
mtcars$cyl <- factor(mtcars$cyl) 
tapply(mtcars$mpg,mtcars$cyl,mean)
```

### Orthogonal Contrasts

```{r}
GM <- c(1/3,1/3,1/3)
H1 <- c(1,-1/2,-1/2)
H2 <- c(0,1,-1)
ex_custom.orthogonal <- rbind(GM,H1,H2)
custom.orthogonal <- calm.encode(ex_custom.orthogonal)
custom.orthogonal
```

```{r}
contrasts(mtcars$cyl) = custom.orthogonal[,-1]
summary(lm(mpg~cyl, mtcars))$coefficients
```

In this custom example, the intercept is equivalent to the mean of the group means. The first parameter represents the difference between the first group and mean of the subsequent groups. The second parameter represents the difference between the second and third groups.

### Nonorthogonal Contrasts

```{r}
GM <- c(1,0,0)
H1 <- c(1,-1/2,-1/2)
H2 <- c(1,-1,0)
ex_custom.nonorthogonal <- rbind(GM,H1,H2)
custom.nonorthogonal <- calm.encode(ex_custom.nonorthogonal)
custom.nonorthogonal
```

```{r}
contrasts(mtcars$cyl) <- custom.nonorthogonal[,-1]
summary(lm(mpg~cyl, mtcars))$coefficients
```

In this custom example, the intercept is equivalent to the mean of the first group. The first parameter represents the difference between the first group and mean of the subsequent groups. The second parameter represents the difference between the first and second groups.
